---
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(results = 'hide')
library("ggplot2")
library("tidyr")
library("dplyr")
```
# Example
It is well-documented in the assigned paper as well as in other literature 
(Agresti & Coull, 1998) that the 95% confidence intervals for one and two 
sample tests derived from Wald tests have poor performance in terms of coverage
rates, with the actual coverage probability being much lower than intended.
This result matches with our own simulations comparing the coverage rates of
different confidence intervals with both one sample and two samples. \
We want to investigate the performance of the adjusted Plus-4 intervals at
other common levels of significance, $\alpha =0.1$ and $\alpha =0.01$ in
comparison with other confidence intervals noted in the article. \
```{r contained}
# checks if a given p is contained in an interval
contained <- function(p, interval) {
  if (p >= interval[1] && p <= interval[2]) {
    return(TRUE)
  } else {
    return(FALSE)
  }
}
```

## Testing One Sample Confidence Intervals at Different $\alpha$ Levels
We simulated 500 datasets from a binomial distribution at different $n$ values
and calculated the coverage rates as a function of the true value of $p$. We
are testing the Wald and the Wilson Score intervals along with the Plus-4.\
For the $\alpha = 0.05$, the results are as expected. The Score and Plus-4
intervals work quite well, and the Wald underperforms for small $n$ values, but
all three has asymptotically similar results as $n$ increases. \
```{r coverage_one}
# calculates the coverage rate of a given interval at a given alpha level by
# simulating datasets from binom(n, p) 
coverage_one <- function(interval, datasets, n, p, alpha) {
  set.seed(1234)
  rsamples <- rbinom(n = datasets, size = n, prob = p) 
  int <- sapply(rsamples, FUN = interval, n = n, alpha = alpha)
  int <- split(int, col(int))
  covered <- sapply(int, FUN = contained, p = p) 
  coverage <-  sum(covered) / datasets 
  return(coverage)
}
```

```{r wald_one_samp}
# calculates the wald interval for one sample of size n from a binomial dist
wald_int_one <- function (sample, n, alpha) {
  mle <- sample/n
  z <- qnorm(1 - alpha/2)
  lower <- mle - z * sqrt((mle)*(1-mle)/n)
  upper <- mle + z * sqrt((mle)*(1-mle)/n)
  c(lower, upper)
}
```

```{r score_one_samp}
# calculates the score interval for a given sample of size n from a binomial dist
score_int_one <- function (sample, n, alpha) {
  mle <- sample/n
  z <- qnorm(1 - alpha/2)
  lower <- (mle+((z^2)/(2*n))-z*sqrt(((mle*(1-mle))/n)+((z^2)/(4*n^2))))/(1+((z^2)/n))
  upper <- (mle+((z^2)/(2*n))+z*sqrt(((mle*(1-mle))/n)+((z^2)/(4*n^2))))/(1+((z^2)/n))
  c(lower, upper)
}
```

```{r plus4_one_samp}
# calculates the plus-4 interval for a given sample of size n from a binomial dist
plus4_int_one <- function (sample, n, alpha) {
  pi <- (sample + 2)/(n + 4)
  z <- qnorm(1 - alpha/2)
  lower <- pi - z*sqrt((pi*(1-pi))/(n+4))
  upper <- pi + z*sqrt((pi*(1-pi))/(n+4))
  c(lower, upper)
}
```

```{r graphs_functions_one_samp}
# functions that graphs simulated coverage rates at a given alpha level
true_p <- seq(0.001, 0.999, 0.001)
simulate_one <- function(datasets, n, alpha) {
  df <- data.frame(
    true_p, 
    wald = sapply(true_p, 
                  FUN = coverage_one,
                  interval = wald_int_one,
                  datasets = datasets,
                  n = n, 
                  alpha = alpha),
    score = sapply(true_p, 
                   FUN = coverage_one,
                   interval = score_int_one,
                   datasets = datasets,
                   n = n, 
                   alpha = alpha),
    plus4 = sapply(true_p, 
                   FUN = coverage_one,
                   interval = plus4_int_one,
                   datasets = datasets,
                   n = n, 
                   alpha = alpha))
  df <- gather(df, key = method, value = coverage, -true_p) %>%
        mutate(n = paste("n =", toString(n)))
  return(df)
}

plot_sims_one <- function(data, alpha, ylim) {
  ggplot(data = data,
       aes(x = true_p, y = coverage, group = method, color = method)) +
  geom_line() +
  geom_hline(yintercept = 1 - alpha) + 
  coord_cartesian(ylim = c(ylim, 1)) +
  labs(
    title = paste("Coverage Rates of Confidence Intervals, a =", toString(alpha)),
    x = "True p", 
    y = "Coverage"
  ) +
  facet_wrap(~n)
}
```

```{r graphs_one_samp_5}
# graphs simulated coverage rates at different alpha = 0.05
alpha <- 0.05
sims <- 500
one_samp_coverage <- rbind(simulate_one(sims, 5, alpha),
                           simulate_one(sims, 20, alpha),
                           simulate_one(sims, 100, alpha),
                           simulate_one(sims, 500, alpha))
one_samp_coverage$n <- factor(one_samp_coverage$n, 
                              levels = c("n = 5","n = 20", "n = 100", "n = 500"))
plot_sims_one(one_samp_coverage, alpha, 0.8)
```
Now we use the same 500 datasets and calculate the coverage rates at $\alpha =0.1$.
For small $n$, the Plus-4 interval now performs slightly worse than the Score
interval.\
```{r graphs_one_samp_10}
# graphs simulated coverage rates at different alpha = 0.1
alpha <- 0.1
one_samp_coverage <- rbind(simulate_one(sims, 5, alpha),
                           simulate_one(sims, 20, alpha),
                           simulate_one(sims, 100, alpha),
                           simulate_one(sims, 500, alpha))
one_samp_coverage$n <- factor(one_samp_coverage$n, 
                              levels = c("n = 5","n = 20", "n = 100", "n = 500"))
plot_sims_one(one_samp_coverage, alpha, 0.75)
```
Again using the same 500 datasets and calculate the coverage rates at $\alpha =0.01$.
For small $n$, the Plus-4 interval now performs noticeably than the Score
interval.\
```{r graphs_one_samp_1}
# graphs simulated coverage rates at different alpha = 0.01
alpha <- 0.01
one_samp_coverage <- rbind(simulate_one(sims, 5, alpha),
                           simulate_one(sims, 20, alpha),
                           simulate_one(sims, 100, alpha),
                           simulate_one(sims, 500, alpha))
one_samp_coverage$n <- factor(one_samp_coverage$n, 
                              levels = c("n = 5","n = 20", "n = 100", "n = 500"))
plot_sims_one(one_samp_coverage, alpha, 0.85)
```

## Testing Two Sample Confidence Intervals at Different $\alpha$ Levels
For two sample confidence intervals, we simulated two groups of 500 datasets
from binomial(n1, p1) and binomial(n2, p2). The confidence interval attempts
to estimate the difference $p_1 - p_2$. We hold $p_2$ fixed and cacluate the
coverage rate as a function of $p_1$. We are testing the Wald and the Hybrid
Score (Newcome Score) intervals along with the Plus-4.\
As expected from the reading, the Score and Plus-4 intervals work quite well
for $\alpha =0.05$ and the Wald underperforms for $p_2$ close to 0 or 1, though
all three intervals performs worse for those extreme $p_2$. \
```{r coverage_two}
# calculates the coverage rate of a given interval at a given alpha level by
# simulating datasets from binom(n1, p1)  and the same number of datasets from
# binom(n2, p2)
coverage_two <- function(interval, datasets, n1, n2, p1, p2, alpha) {
  set.seed(1234)
  rsamples1 <- rbinom(n = datasets, size = n1, prob = p1)
  rsamples2 <- rbinom(n = datasets, size = n2, prob = p2) 
  int <- mapply(sample1 = rsamples1, 
                sample2 = rsamples2,
                FUN = interval, 
                MoreArgs = list(n1 = n1, n2 = n2, alpha = alpha))
  int <- split(int, col(int))
  covered <- sapply(int, FUN = contained, p = (p1 - p2)) 
  coverage <-  sum(covered) / datasets
  return(coverage)
}
```

```{r wald_two_samp}
# calculates the wald interval for one two samples of size n1 and n2 from
# binomial distributions (p1 - p2)
wald_int_two <- function (sample1, sample2, n1, n2, alpha) {
  mle1 <- sample1/n1
  mle2 <- sample2/n2
  z <- qnorm(1 - alpha/2)
  lower <- (mle1 - mle2) - z * sqrt(((mle1*(1-mle1))/n1)+((mle2*(1-mle2))/n2))
  upper <- (mle1 - mle2) + z * sqrt(((mle1*(1-mle1))/n1)+((mle2*(1-mle2))/n2))
  c(lower, upper)
}
```

```{r score_two_samp}
# calculates the hybrid score interval for one two samples of size n1 and n2 
# from binomial distributions (p1 - p2)
score_int_two <- function (sample1, sample2, n1, n2, alpha) {
  mle1 <- sample1/n1
  mle2 <- sample2/n2
  score1 <- score_int_one(sample1, n1, alpha)
  score2 <- score_int_one(sample2, n2, alpha)
  lower <- (mle1 - mle2) - sqrt(((mle1 - score1[1])^2) + ((score2[2] - mle2)^2))
  upper <- (mle1 - mle2) + sqrt(((score1[2] - mle1)^2) + ((mle2 - score2[1])^2))
  c(lower, upper)
}
```

```{r plus4_two_samp}
# calculates the plus-4 interval for one two samples of size n1 and n2 from
# binomial distributions (p1 - p2)
plus4_int_two <- function (sample1, sample2, n1, n2, alpha) {
  p1 <- (sample1 + 1)/(n1 + 2)
  p2 <- (sample2 + 1)/(n2 + 2)
  z <- qnorm(1 - alpha/2)
  lower <- (p1 - p2) - z * sqrt(((p1*(1-p1))/(n1+2)) + ((p2*(1-p2))/(n2+2)))
  upper <- (p1 - p2) + z * sqrt(((p1*(1-p1))/(n1+2)) + ((p2*(1-p2))/(n2+2)))
  c(lower, upper)
}
```


```{r graphs_functions_two_samp}
# functions that graphs simulated coverage rates at a given alpha level
true_p1 <- seq(0.001, 0.999, 0.001)
simulate_two <- function(datasets, n1, n2, p2, alpha) {
  df <- data.frame(
    true_p1, 
    wald = sapply(true_p1, 
                  FUN = coverage_two,
                  interval = wald_int_two, 
                  datasets = datasets, 
                  p2 = p2,
                  n1 = n1, 
                  n2 = n2,
                  alpha = alpha),
    hybrid = sapply(true_p1, 
                    FUN = coverage_two,
                    interval = score_int_two, 
                    datasets = datasets, 
                    p2 = p2,
                    n1 = n1, 
                    n2 = n2,
                    alpha = alpha),
    plus4 = sapply(true_p1, 
                   FUN = coverage_two,
                   interval = plus4_int_two, 
                   datasets = datasets, 
                   p2 = p2,
                   n1 = n1, 
                   n2 = n2,
                   alpha = alpha))
  df <- gather(df, key = method, value = coverage, -true_p1) %>%
        mutate(p2 = paste("p2 =", toString(p2)))
  return(df)
}

plot_sims_two <- function(data, n1, n2, alpha, ylim) {
  ggplot(data = data,
       aes(x = true_p1, y = coverage, group = method, color = method)) +
  geom_line() +
  geom_hline(yintercept = 1 - alpha) + 
  coord_cartesian(ylim = c(ylim, 1)) +
  labs(
    title = paste("Coverage Rates of Confidence Intervals, a = ", 
                  toString(alpha),
                  ", n1 = ",
                  toString(n1),
                  ", n2 = ",
                  toString(n2)),
    x = "True p1", 
    y = "Coverage"
  ) +
  facet_wrap(~p2)
}
```

```{r graphs_two_samp_5}
# graphs simulated coverage rates at different alpha = 0.05, n1 = n2 = 20
alpha <- 0.05
n1 <- 20
n2 <- 20
two_samp_coverage <- rbind(simulate_two(sims, n1, n2, 0.1, alpha),
                           simulate_two(sims, n1, n2, 0.3, alpha),
                           simulate_two(sims, n1, n2, 0.5, alpha),
                           simulate_two(sims, n1, n2, 0.7, alpha))
two_samp_coverage$p2 <- factor(two_samp_coverage$p2, 
                              levels = c("p2 = 0.1","p2 = 0.3", "p2 = 0.5", "p2 = 0.7"))
plot_sims_two(two_samp_coverage, n1, n2, alpha, 0.80)
```
Now we use the same datasets and calculate the coverage rates at $\alpha =0.1$.
Notice that while there is no significant difference between the coverage rates
of the Hybrid Score and the Plus-4, the Plus-4 now underperforms even for $p_2$
near 0.5.\
```{r graphs_two_samp_10}
# graphs simulated coverage rates at different alpha = 0.1, n1 = n2 = 20
alpha <- 0.1
two_samp_coverage <- rbind(simulate_two(sims, n1, n2, 0.1, alpha),
                           simulate_two(sims, n1, n2, 0.3, alpha),
                           simulate_two(sims, n1, n2, 0.5, alpha),
                           simulate_two(sims, n1, n2, 0.7, alpha))
two_samp_coverage$p2 <- factor(two_samp_coverage$p2, 
                              levels = c("p2 = 0.1","p2 = 0.3", "p2 = 0.5", "p2 = 0.7"))
plot_sims_two(two_samp_coverage, n1, n2, alpha, 0.75)
```
Again, the same datasets aer used to calculate the coverage rates at $\alpha =0.1$.
Again, Hybrid Score and the Plus-4 performs well, but now the Plus-4 does not
perform worse for extreme $p_2$ and actually does much better than the Hybrid
for those values.\
```{r graphs_two_samp_1}
# graphs simulated coverage rates at different alpha = 0.01, n1 = n2 = 20
alpha <- 0.01
two_samp_coverage <- rbind(simulate_two(sims, n1, n2, 0.1, alpha),
                           simulate_two(sims, n1, n2, 0.3, alpha),
                           simulate_two(sims, n1, n2, 0.5, alpha),
                           simulate_two(sims, n1, n2, 0.7, alpha))
two_samp_coverage$p2 <- factor(two_samp_coverage$p2, 
                              levels = c("p2 = 0.1","p2 = 0.3", "p2 = 0.5", "p2 = 0.7"))
plot_sims_two(two_samp_coverage, n1, n2, alpha, 0.85)
```
Generally, from the simulations we observed that the adjusted Plus-4 interval
performs well for a confidence level of 95%, underperforms as the confidence
level lowers to 90%, and performs better when the condifence level grows to 99%.\
```{r graphs_two_samp_5_uneven}
# graphs simulated coverage rates at different alpha = 0.05, n1 = n2 = 20
alpha <- 0.05
n1 <- 10
n2 <- 40
two_samp_coverage <- rbind(simulate_two(sims, n1, n2, 0.1, alpha),
                           simulate_two(sims, n1, n2, 0.3, alpha),
                           simulate_two(sims, n1, n2, 0.5, alpha),
                           simulate_two(sims, n1, n2, 0.7, alpha))
two_samp_coverage$p2 <- factor(two_samp_coverage$p2, 
                              levels = c("p2 = 0.1","p2 = 0.3", "p2 = 0.5", "p2 = 0.7"))
plot_sims_two(two_samp_coverage, n1, n2, alpha, 0.80)
```

```{r ref.label='contained', eval=FALSE, echo=TRUE}
```

```{r ref.label='coverage_one', eval=FALSE, echo=TRUE}
```

```{r ref.label='wald_one_samp', eval=FALSE, echo=TRUE}
```

```{r ref.label='score_one_samp', eval=FALSE, echo=TRUE}
```

```{r ref.label='plus4_one_samp', eval=FALSE, echo=TRUE}
```

```{r ref.label='graphs_functions_one_samp', eval=FALSE, echo=TRUE}
```

```{r ref.label='graphs_one_samp_5', eval=FALSE, echo=TRUE}
```

```{r ref.label='graphs_one_samp_10', eval=FALSE, echo=TRUE}
```

```{r ref.label='graphs_one_samp_1', eval=FALSE, echo=TRUE}
```

```{r ref.label='coverage_two', eval=FALSE, echo=TRUE}
```

```{r ref.label='wald_two_samp', eval=FALSE, echo=TRUE}
```

```{r ref.label='score_two_samp', eval=FALSE, echo=TRUE}
```

```{r ref.label='plus4_two_samp', eval=FALSE, echo=TRUE}
```

```{r ref.label='graphs_functions_two_samp', eval=FALSE, echo=TRUE}
```

```{r ref.label='graphs_two_samp_5', eval=FALSE, echo=TRUE}
```

```{r ref.label='graphs_two_samp_10', eval=FALSE, echo=TRUE}
```

```{r ref.label='graphs_two_samp_1', eval=FALSE, echo=TRUE}
```

```{r ref.label='graphs_two_samp_5_uneven', eval=FALSE, echo=TRUE}
```